---
title: "STAT442 Final: Mental Health in Tech Predictions"
output: html_document
author: Kaylee Tringali
date: "December 13th, 2023"
---
### Part I: Preface of Dataset
##### The following dataset comes from a 2014 survey conducted by Open Sourcing Mental Health (formerly OSMI), aiming to measure tech professional's attitudes towards mental health and the frequency of mental health disorders within the field of technology.
```{r}
# load in necessary libraries
library(readr)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(plotly)
library(viridis)
library(ggalt)
library(ggcorrplot)
library(reshape2)
library(relaimpo)

# read in dataset
mental_health1 <- read.csv("/Users/kayleetringali/STAT442 Final/survey_2014.csv")
summary(mental_health1)
```

#### A: preprocessing and cleaning the data
```{r}
# right off the bat, the values for age seem inconsistent
mental_health2 <- mental_health1 %>%
  filter(Age >= 18 & Age <= 65)
summary(mental_health2)

# first, check for missing data
if(sum(is.na(mental_health2)) == 0) {
  print("There is no missing data in our dataset") # displays count of missing values if NA values are present
} else {
  print(paste("There are", sum(is.na(mental_health2)), "missing values in this dataset"))
}

# next, calculate missing values, unique features, and data types for each column
frame <- data.frame(
  Missing_Values = colSums(is.na(mental_health2)),
  Unique_Features = sapply(mental_health2, function(x) length(unique(x))),
  Data_Types = sapply(mental_health2, class)
)
print(frame)

# view unique values in the 'work_interfere' column
unique_values <- unique(mental_health2$work_interfere)

print(unique_values)
```
#### B: Work interference has lots of missing values
```{r}
# define the number of colors needed (assuming 'mental_health2' is your dataset)
num_colors <- length(unique(mental_health2$work_interfere))

# choose a qualitative color palette from RColorBrewer
color_palette <- brewer.pal(num_colors, "Set2")

# create a bar plot with the chosen color palette and without a legend
plot1 <- ggplot(mental_health2, aes(x = work_interfere, fill = work_interfere)) +
  geom_bar() +
  scale_fill_manual(values = color_palette, guide = "none") +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size = 3.5) +
  labs(x = "Work Interference", title = "Work Interference Distribution") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# instead of entirely removing observations with NAs in the work interference and column, fill NAs based on most frequent selection
mental_health3 <- mental_health2 %>%
  mutate(work_interfere = ifelse(is.na(work_interfere),
                                 names(sort(table(work_interfere), decreasing = TRUE)[1]),
                                 work_interfere),
         self_employed = ifelse(is.na(self_employed),
                                names(sort(table(self_employed), decreasing = TRUE)[1]),
                                self_employed))

# create a bar plot for 'work_interfere' column
plot2 <- ggplot(mental_health3, aes(x = work_interfere, fill = work_interfere)) +
  geom_bar() +
  scale_fill_manual(values = color_palette, guide = "none") +
  geom_text(stat = 'count', aes(label = after_stat(count)), vjust = -0.5, size = 3.5) +
  labs(x = "Work Interference", title = "Updated Work Interference Distribution") +
  theme(plot.title = element_text(hjust = 0.5, face = "bold"))

# display the plots
plot1
plot2


```

#### C: Condensing 'Gender' into 3 categories for simplicity and viewing the distribution
```{r}
# replace various gender responses with standardized categories
mental_health2$Gender <- gsub("(?i)^(?=.*\\b(?:male|m|maile|malr|msle|make)\\b).*$", "Male", mental_health2$Gender, perl = TRUE)
mental_health2$Gender <- gsub("(?i)^(?=.*\\b(?:female|f|femake|mal)\\b).*$", "Female", mental_health2$Gender, perl = TRUE)
mental_health2$Gender[mental_health2$Gender == "Guy (-ish) ^_^"] <- "Other"  # replace exact string value

# remove leading/trailing whitespaces
mental_health2$Gender <- trimws(mental_health2$Gender)

# convert non-Male/Female entries to 'Other'
mental_health2$Gender[!(mental_health2$Gender %in% c("Male", "Female"))] <- "Other"

# create a table with counts of each gender category
gender_counts <- table(mental_health2$Gender)

# convert the table into a data frame
gender_counts_df <- as.data.frame(gender_counts)
names(gender_counts_df) <- c("Gender", "Count")

# create a pie chart using plotly
gender_fig <- plot_ly(gender_counts_df, labels = ~Gender, values = ~Count, type = 'pie', hole = 0.4) %>%
  layout(title = "Distribution of Gender Categories", x = 0.5, font = list(size = 13, color = "black", family = "Arial", weight = "bold"))
# show the plot
gender_fig
```

```{r}

# create a table with counts of each age range
age_counts <- table(mental_health2$Age)

# convert the table into a data frame
age_counts_df <- as.data.frame(age_counts)
names(age_counts_df) <- c("Age", "Count")

# create a gradient fill based on the count values
ggplot(age_counts_df, aes(x = Age, y = Count, fill = Count)) +
  geom_bar(stat = "identity") +
  scale_fill_gradient(low = "yellow", high = "red") +  # Define the gradient colors
  labs(title = "Age Distribution", x = "Age", y = "Count") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"))

# create a kernel density plot segmented by treatment categories
ggplot(data = mental_health2, aes(x = Age, fill = treatment)) +
  geom_density(alpha = 0.5) +
  labs(title = "Kernel Density Plot of Age by Treatment", x = "Age", y = "Density") +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5, face = "bold")) +
  scale_fill_manual(name = "Treatment Status",  # Change the legend title
                    values = c("Yes" = "seagreen", "No" = "purple"))  # modify legend colors

```

```{r}

# filter data where treatment is 'Yes'
treated_countries <- mental_health2 %>%
  filter(treatment == 'Yes') %>%
  group_by(Country) %>%
  summarise(count = n()) %>%
  top_n(10, count) %>%
  arrange(desc(count))

# filter data where treatment is 'No'
not_treated_countries <- mental_health2 %>%
  filter(treatment == 'No') %>%
  group_by(Country) %>%
  summarise(count = n()) %>%
  top_n(10, count) %>%
  arrange(desc(count))

# select top 10 treating and not treating countries
treated_countries <- head(treated_countries, 10)
not_treated_countries <- head(not_treated_countries, 10)

# merge the treated and not treated country data
all_countries <- rbind(
  transform(treated_countries, treatment_status = "Treating"),
  transform(not_treated_countries, treatment_status = "Not Treating")
)

# order the merged dataframe by count
all_countries <- all_countries[order(all_countries$count),]

# Create a bar plot for treating and not treating countries
ggplot(all_countries, aes(x = reorder(Country, count), y = count, fill = treatment_status)) +
  geom_bar(stat = "identity", position = "dodge", width = 0.6) +
  labs(title = "Top 10 Countries - Treating vs. Not Treating Mental Health Issues", 
       x = "Country", y = "Frequency") +
  scale_fill_manual(name = "Treatment Status", values = c("Treating" = "seagreen", "Not Treating" = "purple")) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) +
  coord_flip() +
  theme(
    legend.title = element_text(face = NULL),
    plot.title = element_text(hjust = 0.5, face = "bold")
  )

# create a bar plot of anticipated mental health consequences and treatment

# create custom color palette
custom_colors <- c("seagreen", "skyblue")

# create the plot with customizations
ggplot(mental_health2, aes(x = mental_health_consequence, fill = factor(treatment))) +
  geom_bar(position = "dodge") +
  scale_fill_manual(values = custom_colors, name = "Treatment") +
  labs(x = "Mental Health Consequence", y = "Frequency", title = "Frequency of Mental Health Consequence by Treatment") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    legend.title = element_text(face = "bold"),
    legend.position = "top"
  )
```

```{r}

# map categorical variables to numerical values
mental_health3$family_history_num <- ifelse(mental_health3$family_history == "No", 0, 1)
mental_health3$treatment_num <- ifelse(mental_health3$treatment == "No", 0, 1)
mental_health3$self_employed_num <- ifelse(mental_health3$self_employed == "No", 0, 1)
mental_health3$remote_work_num <- ifelse(mental_health3$remote_work == "No", 0, 1)

# convert categorical variables with multiple categories to numerical
mental_health3$benefits_num <- ifelse(mental_health2$benefits == "No", 0, ifelse(mental_health2$benefits == "Yes", 1, 2))
mental_health3$wellness_programs_num <- ifelse(mental_health2$wellness_program == "No", 0, ifelse(mental_health3$wellness_program == "Yes", 1, 2))
mental_health3$seek_help_num <- ifelse(mental_health3$seek_help == "No", 0, ifelse(mental_health3$seek_help == "Yes", 1, 2))
mental_health3$anonymity_num <- ifelse(mental_health3$anonymity == "No", 0, ifelse(mental_health3$anonymity == "Yes", 1, 2))
mental_health3$mental_health_consequence_num <- ifelse(mental_health3$mental_health_consequence == "No", 0, ifelse(mental_health3$mental_health_consequence == "Yes", 1, 2))
mental_health3$phys_health_consequence_num <- ifelse(mental_health3$phys_health_consequence == "No", 0, ifelse(mental_health3$phys_health_consequence == "Yes", 1, 2))

# select numerical columns for correlation analysis
numerical_data <- mental_health3[, c("family_history_num", "treatment_num", "self_employed_num", "remote_work_num", 
                                     "benefits_num", "wellness_programs_num", "seek_help_num", "anonymity_num", 
                                     "mental_health_consequence_num", "phys_health_consequence_num")]

# create correlation matrix
correlation_matrix <- cor(numerical_data)

# plot the correlation heatmap with ggcorrplot
ggcorrplot(correlation_matrix, hc.order = TRUE, 
           type = "lower", lab = TRUE, lab_size = 3,
           method = "circle", outline.color = "white",
           colors = c("blue", "white", "red"), 
           title = "Correlation Heatmap of Numerical Variables")

# create correlation matrix
correlation_matrix <- cor(numerical_data)

# melt the correlation matrix into long format
correlation_melted <- melt(correlation_matrix)

ggplot(correlation_melted, aes(Var1, Var2, fill = value)) +
  geom_tile() +
  geom_text(aes(label = round(value, 2)), size = 3) +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0,
                       limits = c(-1, 1), na.value = "grey50") +
  labs(title = "Correlation Heatmap of Numerical Variables") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold"))  # rotate x-axis labels
```

```{r}
# quantitative analysis

# convert categorical variables to factors
mental_health2$treatment <- as.factor(mental_health2$treatment)
mental_health2$mental_health_consequence <- as.factor(mental_health2$mental_health_consequence)

# convert factors to numeric
mental_health2$treatment_numeric <- as.numeric(mental_health2$treatment) - 1  # Assuming 'No' = 0 and 'Yes' = 1
mental_health2$mental_health_consequence_numeric <- as.numeric(mental_health2$mental_health_consequence) - 1  # Assuming 'No' = 0, 'Maybe' = 1, 'Yes' = 2

# fit the linear regression model
model <- lm(treatment_numeric ~ mental_health_consequence_numeric, data = mental_health2)
summary(model)


```